Species persistence under climate and fishing

### 2014-02-01 Writing/MS.tex
Added explanations for Allee effect, close #20

Previous modeling work examining examining how harvest and range shift intersect has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence. Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions. We built a general model to examine how climate and harvesting interact to affect species persistence. The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.  However, it  ignores many of the complexities present in marine fisheries. We do not include Allee effects, so that even if the population shrank to very low levels it was possible for it to persist over time. However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with Allee effect. We also did not include age structure in our model. The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work. Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.  In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.  By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000} and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.

>>>>

Previous modeling work examining examining how harvest and range shift intersect has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence. Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions. Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.

The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.  However, it  ignores many of the complexities present in marine fisheries. Allee effects, for example, have been suggested as an important factor in fish population dynamics. By leaving out Allee effects, populations which shrank to very low levels were still able for it to persist over time. However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with an Allee effect; i.e. rather than extinction occurring when $n(t)=0$, it would be when $n(t)=\text{the Allee threshold}$. Thus the absolute values of the critical climate velocity and harvesting rates would change, but their relationship to each other and general shape would remain the same.

We also did not include age structure in our model. The effects of both harvesting and climate change may be different across different age classes; including this level of complexity is left for future work. Additionally, we did not include any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.  In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.  By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000} and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.
